---
title: "Spotify Dataset"
author: "Sharjeel Khan"
date: "5/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(fivethirtyeight)
library(ggplot2)
library(dplyr)
```
MATH 2820L: Final Paper (Dataset Analysis)
Names: Sharjeel Khan, [], []

--- 
Instructions:

1, Download the dataset zip (https://www.kaggle.com/yamaerenay/spotify-dataset-19212020-160k-tracks?select=data_by_year_o.csv)
2. Use the getwd() command to find the working and extract the CSV files from the zip into that directory
3. run the read commands to see if RStudio reads the csv files.
--- 
    
```{r}
getwd()
```

First let's take a look at the day by genre file and study it. As we can see, it has all the genres of the music on Spotify and the ratings or qualities associated with the genre such as danceability, instrumentalness, and energy.

```{r}
genres <- read.csv("data_by_genres_o.csv")

head(genres)
dim(genres)

```

Question: What determines a genre's popularity? Here we explores any potential correlation between a genre's popularity with the aforementioned qualities

```{r}
#Here we compare the different aspects of a song to it's popularity.
ggplot(genres, aes(y = popularity, x = danceability )) + 
    geom_point(color = "red", alpha = 0.2)

ggplot(genres, aes(y = popularity, x = energy )) + 
    geom_point(color = "red", alpha = 0.2) 

ggplot(genres, aes(y = popularity , x = instrumentalness )) + 
    geom_point(color = "red", alpha = 0.2) 

ggplot(genres, aes(y = popularity , x = instrumentalness )) + 
    geom_point(color = "red", alpha = 0.2) 

```

From the graphs above, it appears that the studying the music by genre might not be the best way to do it, therefore, let us look at the dataset for individual songs and study it. 

```{r}
songs <- read.csv("data_o.csv")

head(songs)
dim(songs)

```
Scatterplots: 

```{r}
#Here we compare the different aspects of a song to it's popularity.
ggplot(songs, aes(y = popularity, x = danceability )) + 
    geom_point(color = "red", alpha = 0.2) +
    geom_abline(intercept = 0, slope = 0.02, color = "blue")

ggplot(songs, aes(y = popularity, x = energy )) + 
    geom_point(color = "red", alpha = 0.2) +
    geom_abline(intercept = 0, slope = 0.02, color = "blue")

ggplot(songs, aes(y = popularity , x = instrumentalness )) + 
    geom_point(color = "red", alpha = 0.2) +
    geom_abline(intercept = 0, slope = 0.02, color = "blue")

ggplot(songs, aes(y = popularity , x = instrumentalness )) + 
    geom_point(color = "red", alpha = 0.2) +
    geom_abline(intercept = 0, slope = 0.02, color = "blue")

```

